---
title: "Stat 579 - Life Expectancy Homework 1"
date: "9/1/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width=4, fig.height=4)
library(tidyverse)
```

# Life Expectancy in Mexico, Canada and the US 

## Step-by-step plan of work

1. Read the data
2. Focus on values of Canada
3. Visualize the values for Canada
4. See whether we can find the problem
5. ... then we'll see ...

## Reading the data 

```{r}
url <- "https://raw.githubusercontent.com/Stat579-at-ISU/stat579-at-isu.github.io/master/homework/data/gapminder-5060.csv" # nasty looking, but doesn't make any assumptions about data on your machine
life5060 <- read.csv(url)
head(life5060)

```

We read in the gap7080.csv file and check to see how the first few observations look as well.

```{r}
url2 <- "https://raw.githubusercontent.com/Stat579-at-ISU/stat579-at-isu.github.io/master/homework/data/gap7080.csv"
life7080 <- read.csv(url2)
head(life7080)
```


#Combining the data

```{r}
combined <- rbind(life5060,life7080)
head(combined)
tail(combined)
```
We now have a total of 1136 observations of variables. We again take a look at Canada in our newly combined dataset.

## Focus on the values for Canada

```{r}
canada <- combined %>% filter(country == "Canada")
head(canada)

```


## Visualize 

Draw a line for the life expectancy in Canada:
```{r}
canada %>% 
  ggplot(aes(x = year, y = lifeExp)) +
  geom_line()
```

Clearly there is something wrong with the value in 1957. Once again, we have to fix the lifeExp value in the year 1957 for Canada.


## Find the problematic value and fix it

```{r}
canada %>% filter(year == 1957)

canada_fixed <- canada %>% mutate(
  lifeExp = replace(lifeExp, year==1957, 69.96)  
)
```

## Visualize the fixed data

```{r}
canada_fixed %>% ggplot(aes(x = year, y = lifeExp)) +
  geom_line()
```

## ... on to the stretch goal ... 

Add lines for Mexico and the US

```{r}
canada_fixed %>% 
  ggplot(aes(x = year, y = lifeExp, group=country)) + geom_line() +
  geom_line(data = combined %>% 
              filter(country %in% c("United States", "Mexico")))
```

## ... on to the stretch goal ... 

Add lines for Mexico and the US, color lines by country and add a legend

```{r, fig.width = 6}
canada_fixed %>% 
  ggplot(aes(x = year, y = lifeExp, colour=country)) + geom_line() +
  geom_line(data = combined %>% 
              filter(country %in% c("United States", "Mexico")))

```

We now have the Life Expectancy from 1950-1980s for the three countries: Canada, Mexico, and United States in the figure above. We see that life expectancy has increased as the years passed by, with Mexico having the greatest average rate of change out of all the three countries in this time period. We also note that Canada has always had the highest life expectancy in this time frame out of all three countries. 
