---
title: 'Stat 579 - Homework #2 - section B'
author: "Heike Hofmann"
date: "9/10/2019"
output: html_document
---
---
title: 'Stat 579 - Homework #2 - section B'
author: "Heike Hofmann"
date: "9/10/2019"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Homework #2

For this homework we are using the data set `earthquakes` from the classdata package. `earthquake` contains data on every earthquake and earthquake-like event recorded by the USGS (United States Geological Survey) since Jan 1 2019.


Run the following two commands to install the newest version of the package from github:

```{r, eval = FALSE}
library(devtools) # error? then run install.packages("devtools")
devtools::install_github("heike/classdata")
```

Check what is in the data set:
```{r}
library(classdata)
head(earthquakes)
```

1. Download the RMarkdown file with these homework instructions to use as a template for your work.
Make sure to replace "Your Name" in the YAML with your name.

2. Use one of the object inspecting functions discussed in class and interpret the result that you see by applying the function to the earthquakes data. 

3. Answer the following questions (and show your code): how many events were recorded in total? how many of those events are earthquakes? How many of the variables have the word `error` in their name? What is the average magnitude of the observed events?  

4. On July 12, residents of Seattle were rudely awakened by an earthqake of a magnitude of 4.6. What proportion of earthquakes in the data are (roughly) even bigger than the July 12 earthquake in Seattle? Show your code.

5. Visualize the observed geographic location of the events. Extra point for using `ggplot2`.


Due date: the homework is due a week from now on Tuesday 

For the submission: submit your solution in an R Markdown file and (just for insurance) submit the corresponding html/word file with it. 
