---
title: 'Stat 579 - Homework #8 - Section A'
author: "Your Name"
date: "10/24/2019"
output: html_document
---

---
title: 'Stat 579 - Homework #8 - Section A'
author: "Your Name"
date: "10/24/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Urban stream monitoring

For this homework, we will be using the Ames City Urban stream monitoring project available at https://www.cityofames.org/government/departments-divisions-i-z/water-pollution-control/urban-stream-monitoring

1. Download the RMarkdown file with these homework instructions to use as a template for your work.
Make sure to replace "Your Name" in the YAML with your name.
2. Read the information available on the Urban Stream Monitoring project https://www.cityofames.org/government/departments-divisions-i-z/water-pollution-control/urban-stream-monitoring

The table presented on this website can be read into R using the following couple of commands.

```{r, message = FALSE, warning = FALSE}
library(tidyverse)
library(rvest)
url <- "https://www.cityofames.org/government/departments-divisions-i-z/water-pollution-control/urban-stream-monitoring"
doc <- read_html(url)
tables <- doc %>% html_table(fill=TRUE)
ames <- data.frame(tables[[1]])
head(ames)
```
The overarching goal of this homework is to clean the data and get it into a format ready for an analysis.

3. Bring the data into long form. 
4. Measurements of E.Coli are in character format. Convert these values properly into numbers. How many missing values are there? What is the average of the measurements?
5. The `Sample.Date` variable is a mess. Use `separate` to get the information into Month, Day and Year. Make sure that the `Sample.Date` variable is not being removed from the data in this process.
6. Plot measurements by month, colour by creek, and facet by year. Describe the plot in 2-3 sentences.  
7. Introduce a variable `Type` into the dataset that is `fecal coliform bacteria` for measurements before August 14, 2018 and `E.coli` for measurements taken afterwards.
8. `lubridate` is an R package for working with times and dates. Use [RStudio's cheatsheet for lubridate](https://github.com/rstudio/cheatsheets/raw/master/lubridate.pdf) to find a way to convert the variable `Sample.Date` to a date.



Due date: please refer to the website and Canvas for the due date. 

For the submission: submit your solution in an R Markdown file and (just for insurance) submit the corresponding html/word file with it. 
