---
title: 'Stat 579 - Homework #9 - Section A'
author: "Your Name"
date: "10/31/2019"
output: html_document
---

---
title: 'Stat 579 - Homework #9 - Section A'
author: "Your Name"
date: "10/31/2019"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Murder mystery

Github user `sciencificity` has adapted the Murder Mystery by Knight Lab to an R based tidyverse Whodunnit.

1. If you haven't done so yet, install the murder-mystery package by running the following command:
```{r, eval = FALSE}
devtools::install_github("sciencificity/reclues")
```
2. Create an Rmarkdown file with an annotated walkthrough. Make use of `left_join` in at least two instances.
3. Check your solution at http://mystery.knightlab.com/ and follow the additional clues. 
4. Include the solution for the second half of the mystery in your walkthrough.

Due date: please refer to the website and Canvas for the due date. 

For the submission: submit your solution in an R Markdown file and (just for insurance) submit the corresponding html/word file with it. 
