---
title: 'Stat 579 - Homework #4'
author: "Your Name"
date: "9/18/2018"
output: html_document
---

---
title: 'Stat 579 - Homework #4'
author: "Your Name"
date: "9/18/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Subsets and Visualizations: Movie Mojo

For this homework we are using the data set `mojo` from the classdata package. `mojo` constains data on box office revenue for movies based on the website https://www.boxofficemojo.com .


Run the following two commands to install the newest version of the package from github:

```{r, eval = FALSE}
library(devtools) # error? then run install.packages("devtools")
devtools::install_github("heike/classdata")
```

Check what is in the data set:
```{r}
library(classdata)
head(mojo)
```

1. Download the RMarkdown file with these homework instructions to use as a template for your work.
Make sure to replace "Your Name" in the YAML with your name.

2. What is the difference between the variables `Week` and `WeekNo`? Describe in your words. 

3. Use `ggplot2` to plot total gross (`Total Gross`) against week number (`WeekNo`). Facet by Year. Interpret the result. 
Which movie had the highest total gross over the time frame? How many weeks was that movie on rank 1? How long was it in box offices overall?

4. Pick two movies that were in box office some time between 2013 and 2018 and find the corresponding data in the `mojo` data. How does total gross of the two movies compare? Draw a plot and comment on the result. 

5. Hit or Flop? The variable `Budget (in Millions)` contains estimated budget numbers for some movies. For how many movies is this information available (careful! trick question - look at what the function `unique` does)? Studios would like to see their budget returned by the opening weekend. What is the percentage of movies for which that happened? How many movies did not have their budget matched in total gross by the third weekend?
For each of these two questions describe your 'plan of attack', i.e. lay out how you go about finding an answer to the question.
Identify one movie, that did not match its budget by week 3. Plot the incurred loss over time. 


Due date: the homework is due a week from now on Thursday. 

For the submission: submit your solution in an R Markdown file and (just for insurance) submit the corresponding html/word file with it. 
