---
title: 'Stat 579 - Homework #9'
author: "Your Name"
output: html_document
---

---
title: 'Stat 579 - Homework #9'
author: "Your Name"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Murder mystery: Who dunnit?

Our homework this week is an adaption of the Murder Mystery by Knight Lab to an R based tidyverse Whodunnit.

You are given the following clue: 

> A crime has taken place and the detective needs your help. The detective gave you the crime scene report, but you somehow lost it. You vaguely remember that the crime was a murder that occurred sometime on Jan.15, 2018 and that it took place in SQL City. Start by retrieving the corresponding crime scene report from the police department's database.


1. Iinstall the murder-mystery package by running the following command:
```{r, eval = FALSE}
remotes::install_github("sciencificity/reclues")
```
2. The `reclues` package contains  several datasets:

|Data set	|Description	|
|------------------------	|------------------------------------------------	|
| crime_scene_report     	| Crime scene report                             	|
| drivers_license        	| Drivers license                                	|
| facebook_event_checkin 	| Facebook checkin event                          |
| get_fit_now_check_in   	| Fit Now checkin details                        	|
| get_fit_now_member     	| Fit Now member details                         	|
| income                 	| Income                                         	|
| interview              	| Interview                                      	|
| person                 	| Person                                         	|
| solution               	| Solution                                       	|

Start the murder mystery by identifying (and reading) the crime scene report. From there, follow the clues :)
```{r}
library(reclues)
head(crime_scene_report)
```

2. Create an RMarkdown file with an annotated walkthrough. Make use of `left_join` in at least two instances.
3. Check your solution at http://mystery.knightlab.com/ and follow the additional clues. 
4. Include the solution for the second half of the mystery in your walkthrough.

Due date: please refer to the website and Canvas for the due date. 

For the submission: submit your solution in an R Markdown file and (just for insurance) submit the corresponding html/word file with it. 
