---
title: "Stat 579 - Life Expectancy"
date: "8/30/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width=4, fig.height=4)
library(tidyverse)
```

# Life Expectancy in Mexico, Canada and the US 

## Step-by-step plan of work

1. Read the data
2. Focus on values of Canada
3. Visualize the values for Canada
4. See whether we can find the problem
5. ... then we'll see ...

## Reading the data 

```{r}
url5060 <- "https://raw.githubusercontent.com/Stat579-at-ISU/stat579-at-isu.github.io/master/homework/data/gapminder-5060.csv" # nasty looking, but doesn't make any assumptions about data on your machine
life5060 <- read.csv(url5060)
head(life5060)
```

## Focus on the values for Canada

```{r}
canada <- life5060 %>% filter(country == "Canada")
head(canada)
```

## Visualize 

Draw a line for the life expectancy in Canada:
```{r}
canada %>% 
  ggplot(aes(x = year, y = lifeExp)) +
  geom_line()
```

Clearly there is something wrong with the value in 1957.

## Find the problematic value and fix it

```{r}
canada %>% filter(year == 1957)

canada_fixed <- canada %>% mutate(
  lifeExp = replace(lifeExp, year==1957, 69.96)  
)
```

## visualize the fixed data

```{r}
canada_fixed %>% ggplot(aes(x = year, y = lifeExp)) +
  geom_line()
```

## ... on to the stretch goal ... 

Add lines for Mexico and the US

```{r}
canada_fixed %>% 
  ggplot(aes(x = year, y = lifeExp, group=country)) + geom_line() +
  geom_line(data = life5060 %>% 
              filter(country %in% c("United States", "Mexico")))
```

## ... on to the stretch goal ... 

Add lines for Mexico and the US, color lines by country and add a legend

```{r, fig.width = 6}
canada_fixed %>% 
  ggplot(aes(x = year, y = lifeExp, colour=country)) + geom_line() +
  geom_line(data = life5060 %>% 
              filter(country %in% c("United States", "Mexico")))

```


## Adding the second dataset

Here we want to download and store the data for life expectancy from 70 to 80.

```{r}
url7080 <- "https://raw.githubusercontent.com/Stat579-at-ISU/stat579-at-isu.github.io/master/homework/data/gap7080.csv"
life7080 <- read.csv(url7080)

head(life7080)
```

## Merging

Here, we want to combine all datasets to one unified dataset.

``` {r }
life5080 = bind_rows(canada_fixed, life7080, 
                      life5060 %>% filter(country != 'Canada'))

```

## Visualization

Now, we can visualize the data to see the trend about the life expectancy of *US*, *Mexico* and *Canada*.

```{r, fig.width = 6}
life5080 %>% 
  filter(country %in% c("Canada", "United States", "Mexico")) %>%
  ggplot(aes(x=year, y=lifeExp))+
  geom_line(aes(color=country))+
  ggtitle('Life Expectancy')+
  scale_x_continuous(breaks = seq(1952,1987,5))+
  theme(plot.title = element_text(hjust = 0.5))

```

## Comparison

We can compare the results to see how the trend has been changed from 50-60 to 70-80.

```{r, fig.width = 6}
canada_fixed %>% 
  ggplot(aes(x = year, y = lifeExp, colour=country)) + geom_line() +
  geom_line(data = life5060 %>% 
              filter(country %in% c("United States", "Mexico")))+
  geom_line(data=life5080 %>% 
              filter(country %in% c("Canada", "United States", "Mexico")),
            aes(color=country), linetype='dashed')+
  ggtitle('Life Expectancy Comparison')+
  theme(plot.title = element_text(hjust = 0.5))
```


## Conclusions
1. As we can see the Canada has the highest life expectancy, while Mexico has the lowest.

2. The rate of increase for Mexico is bigger than that of US, or Canada. We can clearly see that the differnce is very large at the beginning, but it decreases as time goes on.













