---
title: 'Stat 579 - Homework #2 - section B'
date: "9/10/2019"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Homework #2

For this homework we are using the data set `earthquakes` from the classdata package. `earthquake` contains data on every earthquake and earthquake-like event recorded by the USGS (United States Geological Survey) since Jan 1 2019.


Run the following two commands to install the newest version of the package from github:

```{r, eval = FALSE}
library(devtools) # error? then run install.packages("devtools")
devtools::install_github("heike/classdata")
```

Check what is in the data set:
```{r}
library(classdata)
head(earthquakes)
```


3. Answer the following questions (and show your code): how many events were recorded in total? how many of those events are earthquakes? How many of the variables have the word `error` in their name? What is the average magnitude of the observed events?  
```{r}
summary(earthquakes)
dim(earthquakes)
```
16414 events in total were recorded. 16282 of them were earthquakes. Three variables have the word error in their name. The average magnitude is 3.793.



4. On July 12, residents of Seattle were rudely awakened by an earthqake of a magnitude of 4.6. What proportion of earthquakes in the data are (roughly) even bigger than the July 12 earthquake in Seattle? Show your code.

```{r}
summary(earthquakes$mag)

```
approximately 25% of the earthquakes are stronger than 4.6


5. Visualize the observed geographic location of the events. Extra point for using `ggplot2`.
```{r}
summary(earthquakes$latitude)
summary(earthquakes$longitude)

library(ggplot2)
ggplot(data = earthquakes, aes(x = longitude, y = latitude))+
  geom_point()
```

Due date: the homework is due a week from now on Tuesday 

For the submission: submit your solution in an R Markdown file and (just for insurance) submit the corresponding html/word file with it. 
