---
title: 'Stat 579 - Homework #3 - Section B'
date: "9/10/2018"
output:
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Swiss Banknotes

1. Download the RMarkdown file with these homework instructions to use as a template for your work.
Make sure to replace "Your Name" in the YAML with your name.
2. The R package `alr3` contains a data set called `banknote`, consisting of (physical) measurements on 200 Swiss bank notes, 100 of which are genuine, while the other half is counterfeit. Load this data set (you might have to install the package) using the code below. Also run the cryptic third line - it will make your life a lot easier for the rest of the homework. This turns variable Y explicitly into a factor variable, i.e. makes it discrete. We will discuss this in the course material later in more detail.

```{r, message = FALSE}
# install.packages("alr3")
library(alr3) # if this throws an error of the form 'there is no package called alr3', uncomment the line above, run it once, then comment the line out again and run the code chunk again.
data(banknote)
banknote$Y <- factor(banknote$Y)
```

3. Use one of our object inspecting functions and interpret the result in the data that you see.
```{r}
str(banknote)
```
The `str` command gives that there are 200 observations and 7 variables in the `banknote` dataset. Among the columns, "Length", "Left", "Right", "Bottom", "Top", and "Diagonal" are numeric, while "Y" is factorial. 

4. Use the package `ggplot2` to draw a barchart of Y (0 is genuine, 1 is counterfeit). Map Y to the fill color of the barchart. 
```{r}
library(ggplot2)
library(tidyverse)
banknote %>% ggplot(aes(x = Y, fill = Y)) + 
  geom_bar() +
  theme_light() +
  labs(x = "Y - 0: genuine, 1: counterfeit", y = "Count", title = "Banknote Type vs. Count") +
  theme(plot.title = element_text(hjust = 0.5, face = "bold"))
```

5. Use the package `ggplot2` to draw a histogram of one of the variables in the dataset that shows a distinction between genuine and counterfeit banknotes. Use fill color to show this difference. Choose the binwidth such that there are no gaps in the middle range of the histogram.
```{r}
banknote %>%
  ggplot(aes(x = Diagonal, fill = Y)) + 
  geom_histogram(binwidth = 0.25) +
  theme_light() +
  labs(x = "Diagonal", y = "Count", title = "Diagonal vs. Count") +
  theme(plot.title = element_text(hjust = 0.5, face = "bold"))
```
137.5+(138-137.5)/4 

140.5+(141-140.5)/4

142.5+0.5/4

6. Use the package `ggplot2` to draw a scatterplot of two (continuous) measurements, color by Y. Try to find a pair of measurements that allow you to separate perfectly between genuine and counterfeit banknotes. 
```{r}
banknote %>%
  ggplot(aes(x = Diagonal, y = Right, colour = Y)) + 
  geom_point() +
  theme_light() + 
  labs(x = "Diagonal", y = "Right", title = "Diagonal vs. Right") +
  theme(plot.title = element_text(hjust = 0.5, face = "bold"))
```

7. For each of the three figures above, write a two-three sentence summary, describing the
    a. **structure of the plot**: what type of plot is it? Which variables are mapped to x, to y, and to the (fill) color?
    b. **main message of the plot**: what is your main finding, i.e. what do you want viewers to learn from the plot?
    c. **additional message**: point out anomalies or outliers, if there are any. 

  *Plot in Question 4*
  
  It is a barchart, where variable Y is mapped to x, the number of banknotes is mapped to y, and variable Y is mapped to the fill color. We found that the number of genuine banknotes and the number of counterfeit ones are both 100. There are no outliers shown in the barchart. 
  
  
  *Plot in Question 5*
  
  It is a histogram, where variable Diagonal is mapped to x, the number of banknotes is mapped to y, and variable Y is mapped to the fill color. It can be observed that the counterfeit banknotes tend to have smaller diagonals (from 137.625 to 140.625) than the genuine ones (from 140.625 to 142.625). There are a few outliers. A few banknotes with a diagonal around 139.5 and 140.5 turn out to be genuine.   
  
   
  *Plot in Question 6*
  
  It is a scatterplot, where variable Diagonal is mapped to x, variable Right is mapped to y, and variable Y is mapped to the fill color. There is a clear gap between the genuine banknotes and the counterfeit ones. In general, banknotes with a larger Right/Diagonal ratio tend to be counterfeit while those with a smaller Right/Diagonal ratio tend to be genuine. No anomalies can be observed from the scatterplot. 



Due date: the homework is due a week from now on Tuesday. 

For the submission: submit your solution in an R Markdown file and (just for insurance) submit the corresponding html/word file with it. 
