---
title: 'Stat 579 - Homework #5 - Section B'
date: "10/1/2019"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Behavioral Risk Factor Surveillance System - Again

We are again looking at the data from the
Behavioral Risk Factor Surveillance System (BRFSS).

A subset of the data concentrating on Iowa with records for 2016 is given at
```{r}
url <- "https://raw.githubusercontent.com/Stat579-at-ISU/stat579-at-isu.github.io/master/homework/data/brfss-iowa-2018.csv"
```

The following code reads the data into your R session:
```{r, cache = TRUE, warning = FALSE}
iowa <- read.csv(url)
library(tidyverse)
```

A codebook describing the survey and a listing of all variables is available at [https://www.cdc.gov/brfss/annual_data/2017/pdf/codebook17_llcp-v2-508.pdf]().

For each of the questions, show the code necessary to get to the answer. 
Make sure to also write the answer to the question in a sentence.

1. Download the RMarkdown file with these homework instructions to use as a template for your work.
Make sure to replace "Your Name" in the YAML with your name.
2. Load the dataset into your session and store it in the object `iowa`.

As already seen in the last homework, the coding scheme used for HEIGHT3 is:

| HEIGHT3 value  |  Interpretation    |
|:- |:- |
| 200 - 711 | Height (ft/inches), i.e. 410 is 4 feet, 10 inches	|
| 7777 |	Don't know/Not sure	|
| 9000 - 9998 |	Height (meters/centimeters), where the first 9 indicates that the measurement was metric, 9165 is 1 meter 65 cm |
| 9999	| Refused |
| BLANK |	Not asked or Missing |

The overall goal is to introduce a new variable `height` that corresponds to reported height in centimeters [cm].

3. Sketch out a way to derive a new variable `height` that represents the values in `HEIGHT3` in centimeters and properly encodes any unavailable information as missing values (`NA`).
Implement your transformation. For your convenience: 1 ft equals 30.48 cm, 1 inch equals 2.54 cm. 1 m equals 100 cm.

```{r}
iowa<- iowa %>% mutate(
  height=HEIGHT3,
  height= replace(height, height %in% c(7777, 9999), NA),
  height= ifelse(height > 9000 & height <9998, (height-9000), height),
  height= ifelse(height >= 200 & height <= 711, ((height %/% 100) * 30.48)+ ((height %% 100) * 2.54), height)
)
```
4. How  many missing values does the variable `height` have? 

```{r}
summary(iowa$height)
```
Height has 252 missing values.

5. What is the average height among respondents of the survey? among men? among women? Does the result make sense? Comment.
```{r}
  men_cm <-filter(iowa, SEX1 == 1)
women_cm <-filter(iowa, SEX1 == 2)
summary(men_cm$height)
summary(women_cm$height)
```
The average height for all respondents was 171.23cm. For men, the average height was 178.82cm. For women, the average height was 164.3. 


6. Plot a histogram of `height`, facet by gender such that the histograms are on top of each other. Choose an appropriate bin width and comment on the resulting plot. 

```{r}
iowa$SEX1= factor(iowa$SEX1)
iowa %>%
filter(SEX1 !=7 & SEX1 !=9)%>%
ggplot(aes(x=height, fill=SEX1))+
  geom_histogram(binwidth = 2.5)+
  facet_grid(SEX1 ~ .)+
  scale_fill_manual("Y", values=c("red3", "goldenrod1"))
```


In this graph, you are able to see the distribution of the height values for both men and women represented in centimeters. On avergae, the data is normally distributed by sex, but women are on average shorter than the men. 

7. The variable `WTKG3` has measurements of weight in 10s of grams (i.e. `WTKG3` divided by 100 gives the weight in kilogram). Use `ggplot2` to draw a scatterplot of weight (`WTKG3`) by height (new `height` variable). Facet by gender and comment.

```{r}
iowa$kg= iowa$WTKG3/100
iowa$SEX1= factor(iowa$SEX1)
iowa %>%
filter(SEX1 !=7 & SEX1 !=9)%>%
ggplot(aes(x=height,y= kg, color=SEX1))+
  geom_point()+
  facet_grid(SEX1 ~ .)+
  scale_color_manual("Y", values=c("red3", "goldenrod1"))
```



Through this scatter plot, we are able to see that the men are, on average, heavier and taller than the women. The men have a wider range of weights and heights than the women. 


```
Due date: the homework is due a week from now on Tuesday. 

For the submission: submit your solution in an R Markdown file and (just for insurance) submit the corresponding html/word file with it. 
