---
title: "Stat 579 - Life Expectancy"
date: "8/30/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width=4, fig.height=4)
library(tidyverse)
```

# Life Expectancy in Mexico, Canada and the US 

## Step-by-step plan of work

1. Read the data
2. Focus on values of Canada
3. Visualize the values for Canada
4. See whether we can find the problem
5. ... then we'll see ...

## Reading the data 

```{r}
url <- "https://raw.githubusercontent.com/Stat579-at-ISU/stat579-at-isu.github.io/master/homework/data/gapminder-5060.csv" # nasty looking, but doesn't make any assumptions about data on your machine
url2 <- "https://raw.githubusercontent.com/Stat579-at-ISU/stat579-at-isu.github.io/master/homework/data/gap7080.csv"
life5060 <- read.csv(url)
life7080 <- read.csv(url2)
life50607080 <- rbind(life5060, life7080)
head(life5060)
```

## Focus on the values for Canada

```{r}
canada <- life50607080 %>% filter(country == "Canada")
head(canada)
```

## Visualize 

Draw a line for the life expectancy in Canada:
```{r}
canada %>% 
  ggplot(aes(x = year, y = lifeExp)) +
  geom_line()
```

Clearly there is something wrong with the value in 1957.

## Find the problematic value and fix it

```{r}
canada %>% filter(year == 1957)

canada_fixed <- canada %>% mutate(
  lifeExp = replace(lifeExp, year==1957, 69.96)  
)

```

## visualize the fixed data

```{r}
canada_fixed %>% ggplot(aes(x = year, y = lifeExp)) +
  geom_line()
```

## ... on to the stretch goal ... 

Add lines for Mexico and the US

```{r}
canada_fixed %>% 
  ggplot(aes(x = year, y = lifeExp, group=country)) + geom_line() +
  geom_line(data = life50607080 %>% 
              filter(country %in% c("United States", "Mexico")))
```

## ... on to the stretch goal ... 

Add lines for Mexico and the US, color lines by country and add a legend

```{r, fig.width = 6}
canada_fixed %>% 
  ggplot(aes(x = year, y = lifeExp, colour=country)) + geom_line() +
  geom_line(data = life50607080 %>% 
              filter(country %in% c("United States", "Mexico")))

```

### Some comments about the life expectancy plot above
1. The life expectancies of Canada and US are similar, but significantly longer than that of Mexico for all time. 
2. The life expectancies of all countries kept increasing for all time. And the difference of life expectancies between Mexico and US&Canada became smaller during the studied period.
