---
title: 'Stat 579 - Homework #2'
date: "9/13/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message = FALSE}
library(alr3) 
data(banknote)
banknote$Y <- factor(banknote$Y)
```

3. Use one of our object inspecting functions and interpret the result in the data that you see.

```{r}
str(banknote)
summary(banknote$Y)
```
There are 100 genuine notes and 100 counterfeit notes. 

4. Use the package `ggplot2` to draw a barchart of Y (0 is genuine, 1 is counterfeit). Map Y to the fill color of the barchart. 

```{r}
library(ggplot2)
ggplot(banknote, aes(Y), colour=Y) + 
  geom_bar(aes(fill = Y)) +
  # but could add...
  scale_fill_manual(
   values = c("#F6546A", "#83CCD2"),
    name = "Authenticity", labels = c("Genuine", "Counterfeit"))
```

5. Use the package `ggplot2` to draw a histogram of one of the variables in the dataset that shows a distinction between genuine and counterfeit banknotes. Use fill color to show this difference. Choose the binwidth such that there are no gaps in the middle range of the histogram.

```{r}
ggplot(banknote, aes(x=Diagonal, fill=Y)) +
  geom_histogram(binwidth=0.3) +
  ggtitle("binwidth=0.2") +
  scale_fill_manual(
   values = c("#F6546A", "#83CCD2"),
    name = "Authenticity", labels = c("Genuine", "Counterfeit"))

```

6. Use the package `ggplot2` to draw a scatterplot of two (continuous) measurements, color by Y. Try to find a pair of measurements that allow you to separate perfectly between genuine and counterfeit banknotes. 

```{r}
ggplot(data = banknote, aes(x = Top, 
                           y = Bottom, 
                           colour=Y)) + 
  geom_point() 

```

7. For each of the three figures above, write a two-three sentence summary, describing them

  a. **structure of the plot**: what type of plot is it? Which variables are mapped to x, to y, and to the (fill) color?
    1. The first is a bar chart presenting the frequency of genuine/counterfeit notes. The only variable mapped is "Y", that has the values of 0 = genuine, and 1 = counterfeit. I added colors to the different values to give a distinct look to the plot. 
    2. The second is a histogram showing the variable "Diagonal". It is colored using the variable "Y".
    3. The third plot is a scatterplot of the "Top"(x) and "Bottom" (y) variables. "Y" fills the colors showing the amount of genuine/counterfeit bank notes.  
    
  b. **main message of the plot**: what is your main finding, i.e. what do you want viewers to learn from the plot?
    1. The main message is that there are equal amounts of genuine and counterfeit banknotes in the dataset. 
    2. The plot shows that bank notes with smaller diagonals are more prone to be counterfeit, as oppose to larger diagonals, that are more prone to be genuine. Also, it shows that the range of diagonal measures of counterfeit notes is larger than the genuine notes.
    3. The plot shows that notes with larger top and bottom margins are more prone to be counterfeit. Notes with smaller top and bottom margins are more prone to be genuine. It also shows that counterfeit notes have a more diverse range of format, even with large top margins and small bottom margins. Notes with larger bottom margins are likely to be counterfeit. 
    
  c. **additional message**: point out anomalies or outliers, if there are any. 
    1. Since it is a frequency plot of only one variable with only 2 values, there are no anomalies or outliers. 
    2. The plot shows that, at a certain diagonal measure, which has only occurrences of counterfeit notes, there are more genuine notes than fakes, an anomaly in the distribution.  
    3. The results of the scatterplot surprised me since it shows a lot of different formats for the notes. Also, there are some outliers: one is a genuine note with a large bottom margin and small top margin and one counterfeit note with a large top margin and a small bottom margin. I am not sure how to interpret such results because the shape of such notes looks very unusual. 

