---
title: 'Stat 579 - Homework #4'
date: "9/18/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Subsets and Visualizations: Movie Mojo

For this homework we are using the data set `mojo` from the classdata package. `mojo` constains data on box office revenue for movies based on the website https://www.boxofficemojo.com .


Run the following two commands to install the newest version of the package from github:

```{r, eval = FALSE}
library(devtools) # error? then run install.packages("devtools")
install.packages("devtools")
devtools::install_github("heike/classdata")
```

Check what is in the data set:
```{r}
library(classdata)
head(mojo)
tail(mojo)
```


2. What is the difference between the variables `Week` and `WeekNo`? Describe in your words. 

```{r}
?mojo
```
week is the movie that has been in thearters. weekno is the week number of the year(1 through 52 or 53)


3. Use `ggplot2` to plot total gross (`Total Gross`) against week number (`WeekNo`). Facet by Year. Interpret the result. 

  .from the plot we can see that as the weekno increase, the total gross will increase. the year 2018 has the largest value of total gross than other years.
```{r}
library(ggplot2)
ggplot(mojo, aes(WeekNo, `Total Gross`))+
  geom_point()+
  facet_wrap(~Year)

# Which movie had the highest total gross over the time frame?
library(dplyr)
mojo$Title[which.max(mojo$`Total Gross`)]

# How many weeks was that movie on rank 1?  5
a=filter(mojo, Title=="Star Wars: The Force Awakens", TW==1)
nrow(a)

# How long was it in box offices overall?  24
b=mojo$Week[which(mojo$Title=="Star Wars: The Force Awakens")]
max(b)

```


4. Pick two movies that were in box office some time between 2013 and 2018 and find the corresponding data in the `mojo` data. How does total gross of the two movies compare? Draw a plot and comment on the result. 

from the plot we can see that two movie has the similar total gross in 2018 while "the Meg" has a increasing trend from beginning and "Sreaching" has a higher total gross at the beginning. 
```{r}

a=filter(mojo, Title=="The Meg" | Title=="Searching")
ggplot(a, aes(`Total Gross`, WeekNo))+
  geom_point(aes(colour=factor(Title)))+
  facet_wrap(~Year)
  

```
        

5. Hit or Flop? The variable `Budget (in Million)` contains estimated budget numbers for some movies. For how many movies is this information available (careful! trick question - look at what the function `unique` does)? Studios would like to see their budget returned by the opening weekend. What is the percentage of movies for which that happened? How many movies did not have their budget matched in total gross by the third weekend?
For each of these two questions describe your 'plan of attack', i.e. lay out how you go about finding an answer to the question.
Identify one movie, that did not match its budget by week 3. Plot the incurred loss over time. 

```{r}
# Studios would like to see their budget returned by the opening weekend.???????????????????????? What is the percentage of movies for which that happened? 

mojoNew <- mojo[!is.na(mojo$`Budget (in Million)`),!colnames(mojo) %in% c("Year", "WeekNo")]
mojoNew <- unique(mojoNew)

a=filter(mojoNew, Week==1, `Total Gross`/1000000 > `Budget (in Million)`)$Title
b=filter(mojoNew, Week==1)$Title
ratio1=110/630
ratio1

# How many movies did not have their budget matched in total gross by the third weekend?
c=filter(mojoNew, Week==3, `Total Gross`/1000000 < `Budget (in Million)`)$Title
d=filter(mojoNew, Week==3)$Title
ratio3=320/626
ratio3
#  Identify one movie, that did not match its budget by week 3.Plot the incurred loss over time. 

mojo %>%
  filter( Title == "Mama") %>%
ggplot(aes(x=WeekNo, y=`Weekend Gross`-`Budget (in Million)`))+
  geom_line()+
  facet_wrap(~Year)

```


