---
title: 'Stat 579 - Homework #5'
date: "9/18/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Flying etiquette

FiveThirtyEight is a website founded by Statistician and writer Nate Silver to publish results from  opinion poll analysis, politics, economics, and sports blogging. 
One of the featured articles considers [flying etiquette](https://fivethirtyeight.com/features/airplane-etiquette-recline-seat/). 
This article is based on data collected by FiveThirtyEight and publicly available on github. Use the code below to read in the data from the survey:
```{r}
fly <- read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/flying-etiquette-survey/flying-etiquette.csv")
```

The following couple of lines of code provide a bit of cleanup of the demographic information by reaordering the levels of the  corresponding factor variables. Run this code in your session.

```{r}
fly$Age <- factor(fly$Age, levels=c("18-29", "30-44", "45-60", "> 60", ""))
fly$Household.Income <- factor(fly$Household.Income, levels = c("$0 - $24,999","$25,000 - $49,999", "$50,000 - $99,999", "$100,000 - $149,999", "150000", ""))
fly$Education <- factor(fly$Education, levels = c("Less than high school degree", "High school degree", "Some college or Associate degree", "Bachelor degree",  "Graduate degree", ""))
```

1. Download the RMarkdown file with these homework instructions to use as a template for your work.
Make sure to replace "Your Name" in the YAML with your name.

2. Some people do not travel often by plane. Provide a (visual) breakdown of travel frequency (use variable `How.often.do.you.travel.by.plane.`). Reorder the levels in the variable by travel frequency from least frequent travel to most frequent. Draw a barchart of travel frequency and comment on it. 
Exclude all respondents who never fly from the remainder of the analysis. How many records does the data set have now?
```{r, fig.width = 6}
library(tidyverse)
fly$How.often.do.you.travel.by.plane. <- factor(
  fly$How.often.do.you.travel.by.plane.,
  levels = c("Never", "Once a year or less", "Once a month or less", 
             "A few times per month", "A few times per week", "Every day")
)

ggplot(fly, aes(x = How.often.do.you.travel.by.plane.)) + 
  geom_bar()

```

We can see that there are a large number of respondents who have never flown before, and aside form that, most of the respondents fly once per year or less, with less respondent in each category as the frequency of flights increases.

```{r}
fly <- filter(fly, How.often.do.you.travel.by.plane. != "Never")
nrow(fly)
```

After removing the respondents who never fly, there are 874 records.

3. In the demographic variables (`Education`, `Age`, and `Houshold.Income`), replace all occurrences of the empty string "" by a missing value `NA`. How many responses in each variable do not have any missing values? How many responses have no missing values in any of the three variables? (Hint: think of the function `is.na`)
```{r}
levels(fly$Education)[6] <- NA
levels(fly$Age)[5] <- NA
levels(fly$Household.Income)[6] <- NA

fly %>% 
  filter(!is.na(Education)) %>% 
  nrow()

fly %>% 
  filter(!is.na(Age)) %>% 
  nrow()

fly %>% 
  filter(!is.na(Household.Income)) %>% 
  nrow()

fly %>% 
  filter(!is.na(Education), !is.na(Age), !is.na(Household.Income)) %>% 
  nrow()
```

837 respondents don't have a missing value for Education. 843 don't have a missing value for Age, and 698 don't have a missing value for Household Income. 696 of the 874 responses have no missing values in any of the three variables.

4. Run the command below and interpret the output. What potential purpose can you see for the chart? What might be a problem with the chart? Find at least one purpose and one problem.
```{r, fig.width=8, fig.height=3, message=FALSE}
library(ggplot2)
fly$Education = with(fly, factor(Education, levels = rev(levels(Education))))

ggplot(data = fly, aes(x = 1)) + 
  geom_bar(aes(fill=Education), position="fill") + 
  coord_flip() +
  theme(legend.position="bottom") +
  scale_fill_brewer() + 
  xlab("Ratio") 
```

One purpose might be to see and compare the proportion of respondents within each education level. One problem is that the axis is named incorrectly due to the coordinate flip, so the label "Ratio" should be on the y axis. 

5. Rename the variable `In.general..is.itrude.to.bring.a.baby.on.a.plane.` to `baby.on.plane.`. How many levels does the variable `baby.on.plane` have, and what are these levels? Rename the level labeled "" to "Not answered". Bring the levels of `baby.on.plane` in an order from least rude to most rude. Put the level "Not answered" last. Draw a barchart of variable `baby.on.plane`.  Interpret the result. (This question is very similar to question 2, but preps the data for the next question)

```{r}
fly$baby.on.plane. <- fly$In.general..is.itrude.to.bring.a.baby.on.a.plane.
levels(fly$baby.on.plane.)
levels(fly$baby.on.plane.)[1] <- "Not answered"
fly$baby.on.plane. <- factor(fly$baby.on.plane., 
                             levels = c("No, not at all rude", 
                                        "Yes, somewhat rude",
                                        "Yes, very rude",
                                        "Not answered"))
```

There are four levels of `baby.on.plane.`, they are "", "No, not at all rude," "Yes, somewhat rude," and "Yes, very rude."

```{r, fig.width = 6}
ggplot(fly, aes(x = baby.on.plane.)) + 
  geom_bar()
```

We can see that most people don't think bringing a baby on a plance is rude, while less people think it is very rude than somewhat rude.

6. Investigate the relationship between gender and the variables `Do.you.have.any.children.under.18.` and `baby.on.plane`. How is the attitude towards babies on planes shaped by gender and own children under 18? Find a plot that summarises your findings (use `ggplot2`).
```{r}
fly %>% 
  filter(Gender != "", Do.you.have.any.children.under.18. != "") %>% 
  ggplot(aes(x = Do.you.have.any.children.under.18., fill = baby.on.plane.)) + 
  geom_bar(position = "fill") + 
  facet_wrap(~ Gender) + 
  labs(
    x = "Do you have any children under 18?",
    fill = "Do you think it's rude to take a baby on a plane?"
  )
  
```

We can see that there is larger proportion of men than women who think it is either somewhat or very rude. Additionally, within each gender there is a larger proportion of those who do not have children than those who do who think it is either somewhat or very rude.

Due date: the homework is due a week from now on Thursday. 

For the submission: submit your solution in an R Markdown file and (just for insurance) submit the corresponding html/word file with it. 
