---
title: 'Stat 579 - Homework #6'
date: "10/11/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Another look at the Behavioral Risk Factor Surveillance System 

We are, again, using the data from the Behavioral Risk Factor Surveillance System (BRFSS). Just as a reminder, the BRFSS surveys six individual-level behavioral health risk factors associated with the leading causes of premature mortality and morbidity among adults: 1) cigarette smoking, 2) alcohol use, 3) physical activity, 4) diet, 5) hypertension, and 6) safety belt use. 

A subset of the data concentrating on Iowa with records for 2012 is given at [http://www.hofroe.net/data/iowa-brfss-2012.csv]()

The following code reads the data into your R session:
```{r, warning = FALSE}
iowa <- read.csv("http://www.hofroe.net/data/iowa-brfss-2012.csv")
library(tidyverse)
```

A codebook describing the survey and a listing of all variables is available at [http://www.cdc.gov/brfss/annual_data/2012/pdf/CODEBOOK12_LLCP.pdf]().

For each of the questions, show the code necessary to retrieve the answer. 
Make sure to also write the answer to the question in a sentence.

1. Download the RMarkdown file with these homework instructions to use as a template for your work.
Make sure to replace "Your Name" in the YAML with your name.
2. Load the dataset into your session and store it in the object `iowa`.
3. Check the codebook for an explanation of the variable `DRNKDRI2`. Do a frequency breakdown of the variable `DRNKDRI2` (visually or numerically). Comment.
Introduce a variable `drinkdrive` into the `iowa` data set that encodes 88 as 0, and 77 and 99 as `NA`. <br>
Hint: a combination of `mutate` and `ifelse` might be helpful. <br>
Find the following summaries:
  
    a. What is the average number of times respondents admitted to driving under the influence in the past 30 days?
    b. On how many reports is this average based (exclude missing values)?
    c. Is there a difference in this average between the genders? Is that difference significant? 

```{r}
ggplot(iowa, aes(x = DRNKDRI2)) + 
  geom_bar()

iowa %>% 
  filter(DRNKDRI2 < 77) %>% 
  ggplot(aes(x = DRNKDRI2)) +
  geom_bar()
```

We can see that a majority of the respondents responded 88, but after filtering out those who didn't answer or responded no, we can see that there is a right skew, with most people driving intoxicated only once or twice.

```{r}
iowa <- iowa %>% 
  mutate(
    drunkdrive = ifelse(DRNKDRI2 %in% c(77, 99), NA, 
                        ifelse(DRNKDRI2 == 88, 0, DRNKDRI2))
  )

# a
mean(iowa$drunkdrive, na.rm = TRUE)

# b 
iowa %>% 
  filter(!is.na(drunkdrive)) %>% 
  nrow()

# c
iowa %>% 
  group_by(SEX) %>% 
  summarise(avg = mean(drunkdrive, na.rm = TRUE))

iowa %>% 
  filter(drunkdrive > 0) %>% 
  ggplot(aes(x = drunkdrive)) + 
  geom_bar() + 
  facet_wrap(~ SEX)
```

We can see that the average number of times reported is 0.09304207, based on 3708 respondents. We can also see that the average number of times reported for men is 0.16207091 while it is only 0.02951838. After filtering out those who said 0 and plotting, and considering the averages, we can see that there does seem to be a significant difference in genders.

4. Current smoking status is imputed in the variable `X_SMOKER3` (corresponds to variable `_SMOKER3` in the codebook). Make `X_SMOKER3` a factor.  Relabel levels 1, 2, 3, 4 to `Current Smoker`, `Current Smoker`, `Former Smoker` and `Never Smoked` (yes, `Current Smoker` is repeated on purpose) and level 9 to NA.
Describe the relationship between smoking status and age (use `X_AGE_G` - read up on `_AGE_G` in the codebook) based on an appropriate visualization. 

```{r}
iowa$X_SMOKER3 <- as.factor(iowa$X_SMOKER3)
levels(iowa$X_SMOKER3) <- c("Current Smoker", "Current Smoker", "Former Smoker",
                            "Never Smoked", NA)

ggplot(iowa, aes(x = factor(X_AGE_G), fill = X_SMOKER3)) +
  geom_bar(position = "fill")
```

We can see that as age increases, the proportion who have never smoked decreases. Additionally, the proportion of former smokers increases, which makes sense as you would expect (and hope) more people would quit as they age.

5. What percentage of the population has never smoked? Calculate this percentage by age groups (`X_AGE_G`) and gender (`SEX`). Report also on the number of respondents these percentages are based on.

```{r}
iowa %>% 
  group_by(X_AGE_G, SEX) %>% 
  summarise(
    prop_nvr_smoked = mean(X_SMOKER3 == "Never Smoked", na.rm = T),
    n = n()
  )
```

6. Self assessed health status is used as an indicator in a lot of health andn insurance models. Read up on variable `POORHLTH` in the codebook and recode levels to 'meaningful' numbers (similar to question 3). 
What is the relationship between age and the number of  poor health days? Is this relationship different for men and women? Show plot(s) and describe.

```{r}
iowa <- iowa %>% 
  mutate(
    POORHLTH = ifelse(POORHLTH == 88, 0, 
                      ifelse(POORHLTH %in% c(77, 99), NA, POORHLTH))
  )

iowa %>% 
  ggplot(aes(x = factor(X_AGE_G), y = POORHLTH)) + 
  geom_boxplot()

iowa %>% 
  ggplot(aes(x = factor(X_AGE_G), y = POORHLTH)) + 
  geom_boxplot() + 
  facet_wrap(~ SEX)
```

It seems like as age increases, the more people have a higher number of poor health days. The trend does not seem to differ between sex, but it does seem to be more exaggerated in men than women.

Due date: please refer to the website and Canvas for the due date. 

For the submission: submit your solution in an R Markdown file and (just for insurance) submit the corresponding html/word file with it. 
